Active Learning for SAT Solver Benchmarking {\large (extended and revised version)}

### 2024-05-26 paper_jar/paper.tex
Run Grammarly over text

The approach presented in this paper meets several criteria for benchmarking that are considered desirable (cf.~Table~\ref{tab:requirements}).
Rather than outputting a binary classification, namely whether the new solver is worse than an existing solver or not, we provide a \emph{scoring} function that shows by which margin a solver is worse and how similar its performance is to existing solvers.
In particular, our approach enables \emph{ranking} the new solver among a set of existing solvers.
We show that for ranking, it is not necessary to predict the exact per-instance runtime of the solvers, which would be a more challenging task.
Instead, we work with discretized runtime labels.
We use instance features and known solver runtimes for predicting these labels.
Furthermore, we minimize the \emph{runtime} required for our approach to arrive at its conclusion.
Moreover, we select benchmark instances \emph{non-randomly} and \emph{incrementally}.
In particular, we consider runtime information from already done experiments when choosing the next instance.
By doing so, we can control the properties of the benchmarking approach, such as its required runtime.
Our approach is \emph{scalable} in that it ranks a new solver $\hat{a}$ among any number of known solvers $\mathcal{A}$.
In particular, we only subsample the benchmark once instead of comparing pairwise against each other solver~\cite{MatriconAFSH21}.

@@

We evaluated our approach using the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}, consisting of 5355~instances and runtimes of 28~solvers.
Cross-validation was performed by treating each solver as the new solver once and learning to predict the PAR-2 rank of that solver.
On average, our predictions achieve a ranking accuracy of approximately \SI{92}{\%} with only approximately \SI{10}{\%} of the runtime required to evaluate these solvers on the complete set of instances.
Our entire source code\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}} and experimental data\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking-data}} are available on GitHub.
%JB: data repo not updated since TACAS; does it also contain all the results for the new experiments? as a minimum, it should mention the journal version of the paper (see README of code repo)
%JB: code repo currently has all JAR changes in a branch; probably best to push these changes to "main" for review (in case reviewers have a look at repo but don't switch branches)

@@

This article is an extended and refined version of our conference paper with the title ``Active Learning for SAT Solver Benchmarking''~\cite{fuchs2023active}.
In particular, we significantly extended the analysis of runtime-prediction and runtime-discretization approaches (cf.~Section~\ref{sec:exdesign:disc-pred}), which determined our chosen solver runtime model for instance selection.
Furthermore, we expanded the evaluation of which benchmark instances are selected by our approach (cf.~Section~\ref{sec:eval:instance}).

@@

Benchmarking is of great interest in numerous research areas and represents an active field of research in its own right.
Studies have shown that the compilation of benchmark instances poses a number of challenges.
The use of biased benchmarks can easily lead to fallacious interpretations~\cite{abs-2107-07002}.
Benchmarking also involves a number of interchangeable elements, including the performance measures used, the way the measures are aggregated, and the way missing values are handled.
Questionable research practices could alter these elements a-posteriori in order to meet expectations, thereby skewing the results~\cite{NiesslHWCB22}.

The following sections discusses related work from the areas of static benchmark selection (cf.~Section~\ref{sec:related:static-selection}), algorithm configuration (cf.~Section~\ref{sec:related:algo-configuration}), incremental benchmark selection (cf.~Section~\ref{sec:related:incremental-selection}), and active learning (cf.~Section~\ref{sec:related:active-learning}).
Table~\ref{tab:requirements} compares the most relevant approaches, which all pursue slightly different goals.
In particular, our approach is \emph{not} a general improvement over the others but the only one fully aligned with Definition~\ref{def:new-solver-problem}.

@@

Benchmark selection is essential for solver competitions like the SAT Competition.
In such competitions, the organizers define the rules for composing the benchmarks.
These selection strategies are primarily static, in that they do not depend on particular solvers to distinguish.
Balint et al. provide an overview of benchmark-selection criteria in different solver competitions~\cite{balint2015overview}.
Froleyks et al. describe benchmark selection in recent SAT~competitions~\cite{FroleyksHIJS21}.
Manthey and Möhle propose an approach to remove redundancy from competition benchmarks by considering feature equivalence of formulas~\cite{manthey2016better}.
M{\i}s{\i}r also presents a feature-based approach to reduce benchmarks by using matrix factorization and clustering~\cite{misir2021benchmark}.

Hoos et al.~\cite{HoosKSS13} discuss desirable properties of SAT benchmark instances.
They identify three key selection criteria: instance variety to avoid over-fitting, adapted instance hardness (not too easy but also not too hard), and the avoidance of duplicate instances. 
To filter instances that are too similar, they employ a distance-based approach with the SATzilla features~\cite{XuHHL08,features}.
It should be noted, however, that the approach does not optimize for benchmark \emph{runtime}.
Instead, instances are selected \emph{at random}, with the exception of constraints on instance hardness and feature distance.

@@

Further related work can be found within the field of algorithm configuration~\cite{HoosHL21,Stutzle0P22}, as exemplified by the configuration system SMAC~\cite{HutterHL11}.
Thereby, the goal is to tune SAT solvers for a given sub-domain of problem instances.
While this task differs from our goal, for instance, in that we do not need to navigate the configuration space, there are similarities to our approach as well.
For example, SMAC employs an iterative, model-based selection procedure similar to our approach, but this is for selecting configurations rather than instances.
In contrast, instance selection is made \emph{randomly} in SMAC, i.e., without building a model over instances.
Furthermore, an algorithm configurator cannot be used to \emph{rank} or \emph{score} a new solver relative to others since algorithm configuration is designed to identify the optimal configuration.

@@

Our benchmark selection approach uses active learning.
In passive machine learning, prediction models are trained on datasets with given instance labels (cf.~Fig.~\ref{fig:passive}).
In contrast, active learning (AL) begins with no or little labeled data.
AL then repeatedly selects interesting problem instances for which to acquire labels, gradually improving the prediction model (cf.~Fig.~\ref{fig:activepool}).
AL methods are particularly advantageous when acquiring labels is computationally expensive, such as obtaining solver runtimes.
Without AL methods, it is not evident which instances to label and which not.
In our scenario, the first goal is to maximize the utility of instances to our prediction model, which is measured by the ranking accuracy.
Our second goal is to minimize the cost of label acquisition, which is measured by the instance's predicted runtime.
Consequently, our overall objective is to develop an accurate prediction model without the necessity of labeling each data point.

Rubens et.~al.~\cite{RubensESK15} survey active-learning advances.
While synthesis-based AL methods~\cite{0001AEMN22,GarzonMG22,2019gaal} generate instances for labeling, pool-based methods~\cite{GolbandiKL11,HarpaleY08,KorenBV09} rely on a fixed set of unlabeled instances to sample from.
Recent synthesis-based methods within the field of SAT solving demonstrate the generation of problem instances with desired properties~\cite{0001AEMN22,GarzonMG22}.
However, this objective is distinct from ours.
While those approaches aim to generate instances on which a solver is either effective or ineffective, our objective is to predict whether a solver is effective or ineffective on an \emph{existing} benchmark.
In the latter direction, Volpato and Guangyan employ pool-based AL to train an instance-specific algorithm selector~\cite{volpato2019active}.
However, their objective is not benchmarking a solver's overall performance but recommending the best solver for each SAT instance from a set of available solvers.

\begin{figure}[tbp!]
  \centering
  \begin{tabular}[c]{ccc}
  \begin{subfigure}[b]{0.24\textwidth}
  \centering
  \resizebox{!}{2.75cm}{
  \begin{tikzpicture}
  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
  \node[right] at (-5.9,3.6) {Agent (Solver)};
  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
  \node[right] at (-5.9,2.4) {Training Data};
  \node (v1) at (-2.8,3.4) {};
  \node (v2) at (-2.8,2.8) {};
  \draw[->]  (v1) edge (v2);
  \node[right] at (-2.8,3.1) {$x, y$};
  \node (v3) at (-2.8,2) {};
  \node (v4) at (-2.8,1.4) {};
  \draw[->]  (v3) edge (v4);
  \draw  (-3.1,1.4) rectangle (-2.5,1);
  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
  \node[right] at (-5.9,1.2) {Learning Alg.};
  \node (v5) at (-2.8,1) {};
  \node (v6) at (-2.8,0.4) {};
  \draw[->]  (v5) edge (v6);
  \node[right] at (-5.9,0.25) {Predic. Function};
  \node at (-2.8,0.25) {$f$};
  \end{tikzpicture}
  }
  \caption{Passive.}
  \label{fig:passive}
  \end{subfigure}
  &
  \begin{subfigure}[b]{0.33\textwidth}
  \centering
  \resizebox{!}{2.75cm}{
  \begin{tikzpicture}
  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
  \node[right] at (-5.9,3.6) {Agent (Solver)};
  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
  \node[right] at (-5.9,2.4) {Training Data};
  \node (v1) at (-2.8,3.4) {};
  \node (v2) at (-2.8,2.8) {};
  \draw[->]  (v1) edge (v2);
  \node[right] at (-2.8,3.1) {$x, y$};
  \node (v3) at (-2.8,2) {};
  \node (v4) at (-2.8,1.4) {};
  \draw[->]  (v3) edge (v4);
  \draw  (-3.1,1.4) rectangle (-2.5,1);
  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
  \node[right] at (-5.9,1.2) {Learning Alg.};
  \node (v5) at (-2.8,1) {};
  \node (v6) at (-2.8,0.4) {};
  \draw[->]  (v5) edge (v6);
  \node[right] at (-5.9,0.25) {Predic. Function};
  \node at (-2.8,0.25) {$f$};
  \node[right] at (-1.2,2.4) {Pool};
  \draw[->, densely dashed] (-2.4,1.2) arc (-90:0:0.8);
  \draw[->, densely dashed] (-1.6,2.8) arc (0:90:0.8);
  \node[right] at (-1.6,3.135) {$x, ?$};
  \draw  (-1.6,2.4) ellipse (0.4 and 0.3);
  \draw[fill=black]  (-1.725,2.5) ellipse (0.07 and 0.07);
  \draw[fill=black]  (-1.625,2.25) ellipse (0.07 and 0.07);
  \draw[fill=black]  (-1.425,2.4) ellipse (0.07 and 0.07);
  \end{tikzpicture}
  }
  \caption{Active Pool-based.}
  \label{fig:activepool}
  \end{subfigure}
  &
  \begin{subfigure}[b]{0.33\textwidth}
  \centering
  \resizebox{!}{2.75cm}{
  \begin{tikzpicture}
  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
  \node[right] at (-5.9,3.6) {Agent (Solver)};
  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
  \node[right] at (-5.9,2.4) {Training Data};
  \node (v1) at (-2.8,3.4) {};
  \node (v2) at (-2.8,2.8) {};
  \draw[->]  (v1) edge (v2);
  \node[right] at (-2.8,3.1) {$x, y$};
  \node (v3) at (-2.8,2) {};
  \node (v4) at (-2.8,1.4) {};
  \draw[->]  (v3) edge (v4);
  \draw  (-3.1,1.4) rectangle (-2.5,1);
  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
  \node[right] at (-5.9,1.2) {Learning Alg.};
  \node (v5) at (-2.8,1) {};
  \node (v6) at (-2.8,0.4) {};
  \draw[->]  (v5) edge (v6);
  \node[right] at (-5.9,0.25) {Predic. Function};
  \node at (-2.8,0.25) {$f$};
  \node[right, label={[align=left]Gen.\\Model}] at (-0.75,1.88) {};
  \node[right] at (-1.3,3.135) {$x, ?$};
  \draw[->, densely dashed] (-2.4,1.2) arc (-90:90:1.2);
  \end{tikzpicture}
  }
  \caption{Active Synthesis-based.}
  \label{fig:activesynth}
  \end{subfigure}
  \end{tabular}
  \caption{Types of machine learning (depiction inspired by Rubens~et.al.~\cite{RubensESK15}).}
  \label{fig:learning}
\end{figure}

@@

Algorithm~\ref{algALBenchmark} outlines our benchmarking framework. 
Given a set of solvers~$\mathcal{A}$, instances~$\mathcal{I}$, and runtimes~$r$, we first initialize a prediction model~$\mathcal{M}$ for the new solver $\hat a \not\in \mathcal{A}$ (Line~1).
The prediction model~$\mathcal{M}$ is used to select instances repeatedly (Line~4) for benchmarking the new solver $\hat a$ (Line~5). 
The acquired result is subsequently employed to update the prediction model (Line~7). 
When the stopping criterion is met (Line~3), the benchmarking loop terminates and the final score of $\hat{a}$ is estimated (Line~8). 
Algorithm~\ref{algALBenchmark} returns this score, along with the acquired benchmark instances and runtime measurements.

@@

The model $M$ provides a prediction function $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \mathbb{R}$ for all solvers $\mathcal{\hat A} := \mathcal{A} \cup \lbrace \hat a \rbrace$.
%JB: shouldn't we drop the algorithm as an argument (for $f$ and later for $f'$), since we only train and predict for one solver ($\hat{a}$) anyway? (what would the predictions for the other solvers be -- looking up the exact runtime values from $r$?); instead, could use $\hat{a}$ as a subscript of $f$
This prediction function powers instance selection as described in Section~\ref{sec:main:selection}.
In the update step (Algorithm~\ref{algALBenchmark},~Line~7), the model is trained to predict a transformed version of solver runtime (cf.~Section~\ref{sec:main:model:transformation}) for the new solver~$\hat{a}$, using the features described in Section~\ref{sec:exdesign:data} and the previously acquired runtimes~$\mathcal{R}$.
Each iteration trains a new prediction model from scratch, since the time for running the solver~$\hat{a}$ on one instance~$e$ (Line~5) dominates the training time of the model~$M$ on all acquired instances~$\mathcal{R}$ by a significant margin.
Once the benchmarking loop terminates, the score of the new solver~$\hat{a}$ is predicted and then returned (Lines~8--9).

@@

For the prediction model~$M$, we transform the real-valued runtimes into discrete runtime labels on a per-instance basis.
For each instance $e \in \mathcal{I}$, we use a clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in \mathcal{A} \bigr\}$ to one of $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for the instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_{k-1}$.
Timeouts $\tau$ always form a separate cluster $C_{k}$.
The runtime transformation function $\gamma_k : {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ is then specified as follows:
%
$$\gamma_k(a, e) = j ~\Leftrightarrow~ r(a, e) \in C_j$$
%
Empirical studies on portfolio solvers have demonstrated that discretization is an effective approach in practice~\cite{CollauttiMMO13,NgokoCT19}.
Section~\ref{sec:exdesign:disc-pred} presents a detailed analysis of the runtime transformation in our scenario.
In particular, the analysis demonstrates that our benchmarking approach achieves higher ranking accuracy when working with discrete runtime labels rather than raw runtimes.

@@

The uncertainty sampling strategy selects the instance that is closest to the model's decision boundary.
This is achieved by selecting the instance $e \in \mathcal{I} \setminus \tilde{\mathcal{I}}$ that minimizes the certainty~$U(e)$, which is specified by the following equation:
%
\begin{equation*}
  \operatorname{U}(e) := \left\lvert \frac{1}{k} - \max_{n \in \left\lbrace 1, \dots, k \right\rbrace} f'\!\left(\hat{a}, e\right)_{n} \right\rvert
\end{equation*}

@@

In addition, we retrieve instance family information\footnote{\url{https://benchmark-database.de/getdatabase/meta_db}} to evaluate the composition of our sampled benchmarks.
Instance families consist of instances from the same application domain, such as planning, cryptography, etc., and are a valuable tool for analyzing solver performance.
%
For hyperparameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instance family.
All instance families that are too \emph{small}, i.e., \SI{10}{\%} of them correspond to less than one instance, are put into one meta-family for stratification.
%JB: i.e., families with less than 10 instances? could be phrased more directly
This \emph{tuning dataset} allows a more extensive exploration of the hyperparameter space.

@@

One of the most important parts of our framework is the solver runtime model (cf.~Section~\ref{sec:main:model}).
In this section, we detail how clustering solver runtimes can lead to informative runtime labels, and which machine learning models are best at predicting them.
To this end, we analyze prediction performance for estimating runtimes, timeouts, and discretized runtimes.

>>>>

The approach presented in this paper meets several criteria for benchmarking that are considered desirable (cf.~Table~\ref{tab:requirements}).
Rather than outputting a binary classification, namely whether the new solver is worse than an existing solver or not, we provide a \emph{scoring} function that shows by which margin a solver is worse and how similar its performance is to existing solvers.
In particular, our approach enables \emph{ranking} the new solver among a set of existing solvers.
We show that predicting the exact per-instance runtime of the solvers, which would be a more challenging task, is not necessary for ranking.
Instead, we work with discretized runtime labels.
We use instance features and known solver runtimes to predict these labels.
Furthermore, we minimize the \emph{runtime} required for our approach to arrive at its conclusion.
Moreover, we select benchmark instances \emph{non-randomly} and \emph{incrementally}.
In particular, we consider runtime information from experiments already done when choosing the next instance.
By doing so, we can control the properties of the benchmarking approach, such as its required runtime.
Our approach is \emph{scalable} in that it ranks a new solver $\hat{a}$ among any number of known solvers $\mathcal{A}$.
In particular, we only subsample the benchmark once instead of comparing pairwise against each other solver~\cite{MatriconAFSH21}.

@@

We evaluated our approach using the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}, consisting of 5355~instances and runtimes of 28~solvers.
Cross-validation was performed by treating each solver as the new solver once and learning to predict its PAR-2 rank.
On average, our predictions achieve a ranking accuracy of approximately \SI{92}{\%} with only approximately \SI{10}{\%} of the runtime required to evaluate these solvers on the complete set of instances.
Our entire source code\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}} and experimental data\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking-data}} are available on GitHub.
%JB: data repo not updated since TACAS; does it also contain all the results for the new experiments? as a minimum, it should mention the journal version of the paper (see README of code repo)
%JB: code repo currently has all JAR changes in a branch; probably best to push these changes to "main" for review (in case reviewers have a look at repo but don't switch branches)

@@

This article is an extended and refined version of our conference paper titled ``Active Learning for SAT Solver Benchmarking''~\cite{fuchs2023active}.
In particular, we significantly extended the analysis of runtime-prediction and runtime-discretization approaches (cf.~Section~\ref{sec:exdesign:disc-pred}), which determined our chosen solver runtime model for instance selection.
Furthermore, we expanded the evaluation of which benchmark instances are selected by our approach (cf.~Section~\ref{sec:eval:instance}).

@@

Benchmarking is of great interest in numerous research areas and represents an active field of research in its own right.
Studies have shown that the compilation of benchmark instances poses several challenges.
Using biased benchmarks can easily lead to fallacious interpretations~\cite{abs-2107-07002}.
Benchmarking also involves several interchangeable elements, including the performance measures used, how the measures are aggregated, and how missing values are handled.
Questionable research practices could alter these elements a-posteriori to meet expectations, thereby skewing the results~\cite{NiesslHWCB22}.

The following sections discuss related work from the areas of static benchmark selection (cf.~Section~\ref{sec:related:static-selection}), algorithm configuration (cf.~Section~\ref{sec:related:algo-configuration}), incremental benchmark selection (cf.~Section~\ref{sec:related:incremental-selection}), and active learning (cf.~Section~\ref{sec:related:active-learning}).
Table~\ref{tab:requirements} compares the most relevant approaches, which all pursue slightly different goals.
In particular, our approach is \emph{not} a general improvement over the others but the only one fully aligned with Definition~\ref{def:new-solver-problem}.

@@

Benchmark selection is essential for solver competitions like the SAT Competition.
In such competitions, the organizers define the rules for composing the benchmarks.
These selection strategies are primarily static since they do not depend on particular solvers to distinguish.
Balint et al. provide an overview of benchmark-selection criteria in different solver competitions~\cite{balint2015overview}.
Froleyks et al. describe benchmark selection in recent SAT~competitions~\cite{FroleyksHIJS21}.
Manthey and Möhle propose an approach to remove redundancy from competition benchmarks by considering feature equivalence of formulas~\cite{manthey2016better}.
M{\i}s{\i}r also presents a feature-based approach to reduce benchmarks by using matrix factorization and clustering~\cite{misir2021benchmark}.

Hoos et al.~\cite{HoosKSS13} discuss desirable properties of SAT benchmark instances.
They identify three key selection criteria: instance variety to avoid over-fitting, adapted instance hardness (neither too easy nor too hard), and avoiding duplicate instances. 
To filter instances that are too similar, they employ a distance-based approach with the SATzilla features~\cite{XuHHL08,features}.
It should be noted, however, that the approach does not optimize for benchmark \emph{runtime}.
Instead, instances are selected \emph{at random} except considering constraints on instance hardness and feature distance.

@@

Further related work can be found within the field of algorithm configuration~\cite{HoosHL21,Stutzle0P22}, as exemplified by the configuration system SMAC~\cite{HutterHL11}.
Thereby, the goal is to tune SAT solvers for a given sub-domain of problem instances.
While this task differs from our goal, for instance, in that we do not need to navigate the configuration space, there are similarities to our approach as well.
For example, SMAC employs an iterative, model-based selection procedure similar to our approach, but this is for selecting configurations rather than instances.
In contrast, instance selection is made \emph{randomly} in SMAC, i.e., without building a model over instances.
Furthermore, since algorithm configuration is designed to identify the optimal configuration, an algorithm configurator cannot be used to \emph{rank} or \emph{score} a new solver relative to others.

@@

Our benchmark selection approach uses active learning.
In passive machine learning, prediction models are trained on datasets with given instance labels (cf.~Fig.~\ref{fig:passive}).
In contrast, active learning (AL) begins with little or no labeled data.
AL then repeatedly selects interesting problem instances for which to acquire labels, gradually improving the prediction model (cf.~Fig.~\ref{fig:activepool}).
AL methods are particularly advantageous when acquiring labels is computationally expensive, such as obtaining solver runtimes.
Without AL methods, it is not evident which instances to label and which not.
In our scenario, the first goal is to maximize the utility of instances to our prediction model, which is measured by the ranking accuracy.
Our second goal is to minimize the cost of label acquisition, which is measured by the instance's predicted runtime.
Consequently, our overall objective is to develop an accurate prediction model without labeling each data point.

Rubens et~al.~\cite{RubensESK15} survey active-learning advances.
While synthesis-based AL methods~\cite{0001AEMN22,GarzonMG22,2019gaal} generate instances for labeling, pool-based methods~\cite{GolbandiKL11,HarpaleY08,KorenBV09} rely on a fixed set of unlabeled instances to sample from.
Recent synthesis-based methods within the field of SAT solving demonstrate the generation of problem instances with desired properties~\cite{0001AEMN22,GarzonMG22}.
However, this objective is distinct from ours.
While those approaches aim to generate instances on which a solver is either effective or ineffective, our objective is to predict whether a solver is effective or ineffective on an \emph{existing} benchmark.
In the latter direction, Volpato and Guangyan employ pool-based AL to train an instance-specific algorithm selector~\cite{volpato2019active}.
However, their objective is not to benchmark a solver's overall performance but to recommend the best solver for each SAT instance from a set of available solvers.

\begin{figure}[tbp!]
  \centering
  \begin{tabular}[c]{ccc}
  \begin{subfigure}[b]{0.24\textwidth}
  \centering
  \resizebox{!}{2.75cm}{
  \begin{tikzpicture}
  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
  \node[right] at (-5.9,3.6) {Agent (Solver)};
  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
  \node[right] at (-5.9,2.4) {Training Data};
  \node (v1) at (-2.8,3.4) {};
  \node (v2) at (-2.8,2.8) {};
  \draw[->]  (v1) edge (v2);
  \node[right] at (-2.8,3.1) {$x, y$};
  \node (v3) at (-2.8,2) {};
  \node (v4) at (-2.8,1.4) {};
  \draw[->]  (v3) edge (v4);
  \draw  (-3.1,1.4) rectangle (-2.5,1);
  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
  \node[right] at (-5.9,1.2) {Learning Alg.};
  \node (v5) at (-2.8,1) {};
  \node (v6) at (-2.8,0.4) {};
  \draw[->]  (v5) edge (v6);
  \node[right] at (-5.9,0.25) {Predic. Function};
  \node at (-2.8,0.25) {$f$};
  \end{tikzpicture}
  }
  \caption{Passive.}
  \label{fig:passive}
  \end{subfigure}
  &
  \begin{subfigure}[b]{0.33\textwidth}
  \centering
  \resizebox{!}{2.75cm}{
  \begin{tikzpicture}
  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
  \node[right] at (-5.9,3.6) {Agent (Solver)};
  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
  \node[right] at (-5.9,2.4) {Training Data};
  \node (v1) at (-2.8,3.4) {};
  \node (v2) at (-2.8,2.8) {};
  \draw[->]  (v1) edge (v2);
  \node[right] at (-2.8,3.1) {$x, y$};
  \node (v3) at (-2.8,2) {};
  \node (v4) at (-2.8,1.4) {};
  \draw[->]  (v3) edge (v4);
  \draw  (-3.1,1.4) rectangle (-2.5,1);
  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
  \node[right] at (-5.9,1.2) {Learning Alg.};
  \node (v5) at (-2.8,1) {};
  \node (v6) at (-2.8,0.4) {};
  \draw[->]  (v5) edge (v6);
  \node[right] at (-5.9,0.25) {Predic. Function};
  \node at (-2.8,0.25) {$f$};
  \node[right] at (-1.2,2.4) {Pool};
  \draw[->, densely dashed] (-2.4,1.2) arc (-90:0:0.8);
  \draw[->, densely dashed] (-1.6,2.8) arc (0:90:0.8);
  \node[right] at (-1.6,3.135) {$x, ?$};
  \draw  (-1.6,2.4) ellipse (0.4 and 0.3);
  \draw[fill=black]  (-1.725,2.5) ellipse (0.07 and 0.07);
  \draw[fill=black]  (-1.625,2.25) ellipse (0.07 and 0.07);
  \draw[fill=black]  (-1.425,2.4) ellipse (0.07 and 0.07);
  \end{tikzpicture}
  }
  \caption{Active Pool-based.}
  \label{fig:activepool}
  \end{subfigure}
  &
  \begin{subfigure}[b]{0.33\textwidth}
  \centering
  \resizebox{!}{2.75cm}{
  \begin{tikzpicture}
  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
  \node[right] at (-5.9,3.6) {Agent (Solver)};
  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
  \node[right] at (-5.9,2.4) {Training Data};
  \node (v1) at (-2.8,3.4) {};
  \node (v2) at (-2.8,2.8) {};
  \draw[->]  (v1) edge (v2);
  \node[right] at (-2.8,3.1) {$x, y$};
  \node (v3) at (-2.8,2) {};
  \node (v4) at (-2.8,1.4) {};
  \draw[->]  (v3) edge (v4);
  \draw  (-3.1,1.4) rectangle (-2.5,1);
  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
  \node[right] at (-5.9,1.2) {Learning Alg.};
  \node (v5) at (-2.8,1) {};
  \node (v6) at (-2.8,0.4) {};
  \draw[->]  (v5) edge (v6);
  \node[right] at (-5.9,0.25) {Predic. Function};
  \node at (-2.8,0.25) {$f$};
  \node[right, label={[align=left]Gen.\\Model}] at (-0.75,1.88) {};
  \node[right] at (-1.3,3.135) {$x, ?$};
  \draw[->, densely dashed] (-2.4,1.2) arc (-90:90:1.2);
  \end{tikzpicture}
  }
  \caption{Active Synthesis-based.}
  \label{fig:activesynth}
  \end{subfigure}
  \end{tabular}
  \caption{Types of machine learning (depiction inspired by Rubens~et~al.~\cite{RubensESK15}).}
  \label{fig:learning}
\end{figure}

@@

Algorithm~\ref{algALBenchmark} outlines our benchmarking framework. 
Given a set of solvers~$\mathcal{A}$, instances~$\mathcal{I}$, and runtimes~$r$, we first initialize a prediction model~$\mathcal{M}$ for the new solver $\hat a \not\in \mathcal{A}$ (Line~1).
The prediction model~$\mathcal{M}$ is used to select instances repeatedly (Line~4) for benchmarking the new solver $\hat a$ (Line~5). 
The acquired result is subsequently employed to update the prediction model (Line~7). 
When the stopping criterion is met (Line~3), the benchmarking loop terminates, and the final score of $\hat{a}$ is estimated (Line~8). 
Algorithm~\ref{algALBenchmark} returns this score, along with the acquired benchmark instances and runtime measurements.

@@

The model $M$ provides a prediction function $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \mathbb{R}$ for all solvers $\mathcal{\hat A} := \mathcal{A} \cup \lbrace \hat a \rbrace$.
%JB: shouldn't we drop the algorithm as an argument (for $f$ and later for $f'$), since we only train and predict for one solver ($\hat{a}$) anyway? (what would the predictions for the other solvers be -- looking up the exact runtime values from $r$?); instead, could use $\hat{a}$ as a subscript of $f$
This prediction function powers instance selection as described in Section~\ref{sec:main:selection}.
In the update step (Algorithm~\ref{algALBenchmark},~Line~7), the model is trained to predict a transformed version of solver runtime (cf.~Section~\ref{sec:main:model:transformation}) for the new solver~$\hat{a}$, using the features described in Section~\ref{sec:exdesign:data} and the previously acquired runtimes~$\mathcal{R}$.
Each iteration trains a new prediction model from scratch since the time for running the solver~$\hat{a}$ on one instance~$e$ (Line~5) dominates the training time of the model~$M$ on all acquired instances~$\mathcal{R}$ by a significant margin.
Once the benchmarking loop terminates, the score of the new solver~$\hat{a}$ is predicted and then returned (Lines~8--9).

@@

For the prediction model~$M$, we transform the real-valued runtimes into discrete runtime labels on a per-instance basis.
For each instance $e \in \mathcal{I}$, we use a clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in \mathcal{A} \bigr\}$ to one of $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for the benchmark instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_{k-1}$.
Timeouts $\tau$ always form a separate cluster $C_{k}$.
The runtime transformation function $\gamma_k : {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ is then specified as follows:
%
$$\gamma_k(a, e) = j ~\Leftrightarrow~ r(a, e) \in C_j$$
%
Empirical studies on portfolio solvers have demonstrated that discretization is an effective approach in practice~\cite{CollauttiMMO13,NgokoCT19}.
Section~\ref{sec:exdesign:disc-pred} presents a detailed analysis of the runtime transformation in our scenario.
In particular, the analysis demonstrates that our benchmarking approach achieves higher ranking accuracy when working with discrete runtime labels rather than raw runtimes.

@@

The uncertainty sampling strategy selects the instance closest to the model's decision boundary.
This is achieved by selecting the instance $e \in \mathcal{I} \setminus \tilde{\mathcal{I}}$ that minimizes the certainty~$U(e)$, which is specified by the following equation:
%
\begin{equation*}
  \operatorname{U}(e) := \left\lvert \frac{1}{k} - \max_{n \in \left\lbrace 1, \dots, k \right\rbrace} f'\!\left(\hat{a}, e\right)_{n} \right\rvert
\end{equation*}

@@

In addition, we retrieve instance family information\footnote{\url{https://benchmark-database.de/getdatabase/meta_db}} to evaluate the composition of our sampled benchmarks.
Instance families consist of instances from the same application domain, such as planning, cryptography, etc., and are valuable for analyzing solver performance.
%
For hyperparameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instance family.
All instance families that are too \emph{small}, i.e., \SI{10}{\%} of them correspond to less than one instance, are put into one meta-family for stratification.
%JB: i.e., families with less than 10 instances? could be phrased more directly
This \emph{tuning dataset} allows a more extensive exploration of the hyperparameter space.

@@

One of the most important parts of our framework is the solver runtime model (cf.~Section~\ref{sec:main:model}).
In this section, we detail how clustering solver runtimes can lead to informative runtime labels and which machine learning models are best at predicting them.
To this end, we analyze prediction performance for estimating runtimes, timeouts, and discretized runtimes.
